---
title: "Untitled"
author: "Katherine Pulham"
date: "2/24/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggfortify)
library(gridExtra)
```

I want to model the model wide average maximum temperature as a time series. I've used hyper_tibble() to create a csv for it:

```{r}
maxtfile <- "ERA-Interim_averages/average-Max-Temperature.csv"
maxt <- read_csv(maxtfile)
```

I clean this up really quickly

```{r}
dvec <- character(length=length(maxt$avg.max.temp))
for (i in 1:length(maxt$avg.max.temp)){
  dvec[i] <- paste(maxt$year[i],maxt$month[i],maxt$day[i],sep="-")
}
maxt <- maxt %>%
  mutate(date=lubridate::ymd(dvec),maxtemp=avg.max.temp)
maxt <- select(.data=maxt,date,maxtemp)
head(maxt)
remove(dvec)
```

Let's start by plotting this.

```{r}
ggplot(data=maxt, aes(x=date,y=maxtemp))+
  geom_line()+
  ylim(c(250,330))+
  ggtitle("Daily Maximum Temperature")+
  ylab("Maximum Temperature (degrees Kelvin)")+
  xlab("Date")
```

This appears to be a sine wave, maybe with a slight trend to it. Since the oscillations have to have a year long period, I'll be fitting the model:

$$
Y=\beta_0+\beta_1X+\alpha \cdot sin\left(\frac{2\pi X}{365.25}+\omega\right)
$$

This is not linear in terms of the parameters $\underset{\sim}{\beta}$, we must use a non linear least squares estimation scheme.

```{r}
#creating a time variable for days since 1979-1-1 and the temp
t <- 1:length(maxt$date)
temp <- maxt$maxtemp

#fitting the nls model
fit <- nls(temp ~ beta0 + beta1*t + alpha*sin((2*pi/365.25)*t + omega),
             start = list(beta0=290, alpha=10, omega=4.9, beta1=.000000000001))
summary(fit)
```

We see a very small, but positive slope term for the trend on the data. Let's plot this fitted curve:

```{r}
#set parameters
parameters <- summary(fit)$coef[,1]
b0    <- parameters[1]
b1    <- parameters[4]
alpha <- parameters[2]
omega <- parameters[3]

#define prediction function
pred <- function(t){
  return(b0 + b1*t + alpha*sin((2*pi*t/365.25)+omega))
}


#produce predicted values and add to maxt
yhat <- pred(t)

#plot the data and the model
ggplot(data=NULL, aes(x=t,y=yhat))+geom_line(col="green")+
  geom_line(data=NULL, aes(x=t,y=temp))+
  xlim(1,1000)
```

Now we can extract the residuals from this model and examine them as a time series, plotting the series, the acf, and the pacf:

```{r}
rdf <- tibble(
  t = (1:length(resid(fit))),
  residuals = resid(fit)
)
p0 <- ggplot(data=rdf)+geom_line(aes(x=t,y=residuals))
p1 <- autoplot(acf(rdf$residuals,plot=F))
p2 <- autoplot(pacf(rdf$residuals,plot=F))
grid.arrange(p0,p1,p2, nrow = 3) 
```

